Envelope Estimation using Geometric Properties of a Discrete Real Signal

### 2021-05-11 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
revising

\begin{abstract} 
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Envelope detection techniques have applications in medicine, sound classification, sound synthesis, seismology, and speech recognition. Nevertheless, a general approach to envelope detection doesn't exist. Most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation. To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning. The method draws inspiration from geometric concepts; the discrete signal is mapped to a set of points in the Cartesian plane, where a new measure of discrete curvature is used to identify the samples that are part of the envelope. Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave. The algorithm compares favorably with classic envelope detection techniques based on smoothing, filtering, and the Hilbert Transform, besides being physically plausible. We provide visualizations of the envelope extracted for various real-world signals with very diverse characteristics, such as voice, spoken and sang, and pitched and unpitched musical instruments, and discuss some approaches to objectively assess the quality of the obtained envelopes. An optimized implementation is available via the Python Package Index (PyPI); interactive visualizations and source code are available at the accompanying repository.
\end{abstract}

@@

\section{Introduction}
Envelope detection, also known as demodulation, is ubiquitous in analog and digital signal processing (DSP) \cite{2011CaetanoImproved}. Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for particular settings, in general involving narrowband signals with moderate noise content.

Those constraints exclude most physical signals, as is the case of recorded sound, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}.

@@

The temporal envelope also helps to impart emotion and identity to the human voice \cite{2018ZhuContributions} and music: Concert halls with envelope preserving characteristics are deemed more pleasant \cite{2011LokkiEngaging}.

Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting prior knowledge of the waves, are prevalent, such as the one presented by \cite{2014YangFast} for the distributed monitoring of fiber optic or the one formulated by \cite{2018AssefModeling} in the context of medical ultrasound imaging.

The problem of envelope detection of a discrete signal, normally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.

Considering the standard graphic representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents the amplitude, one can think of each sample as a point in a plane. Provided a coherent mapping from this space to the Cartesian plane is available, the original DSP problem can be transformed into the geometric problem of finding the shape of this set of points.

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is by using the concept of a convex hull, as defined in \cite{2008BergComputational}. The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $. Convex hulls are unique, in the sense that a given set of points define one and only one convex hull.

@@

Fig. \ref{fig:AlphaShape} illustrates an alpha shape, for $ \alpha=10 $, of a set of points. Note that two points of the set lie at the circle illustrated, but the open disc defined by it contains no points of the set. An analogous construction is possible for every point at the frontier.

Conversely, any circle of radius $ \alpha=10 $ that contains at least one of the interior defines an open disc that also contains at least another point of the set P.

@@

Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property: for a given set of points and a given radius, only one alpha shape exists. When the radius $ \alpha $ of the circle tends to infinity, alpha shapes, also known as concave hulls, tend to the convex hulls.

Consider, for example, the artificially generated discrete wave in Fig. \ref{fig:Hulls} below, where the links between consecutive samples were de-emphasized to highlight the “set of points” nature of a signal: two alpha shapes, with different values for $ \alpha $, and the convex hull are illustrated. As $ \alpha $ increases, the contour of the set of points becomes smoother, and closer to the convex hull.

@@

The alpha shapes theory is used in areas such as detection of features in images \cite{2016VarytimidisAlpha}, reconstruction of surfaces from a cloud of points \cite{2015WuAutomated} and spectroscopy \cite{2019XuModeling}.

@@

Other steps in the direction of translating geometric algorithms to the context of envelope detection were made by \cite{2015YangSkeleton} via an algorithm based on the construction of a skeleton underlying the digital wave of interest, and also via the direct translation of computer vision methods to the task of envelope detection \cite{2015YangRepresenting}.

@@

We start with a brief background of envelope detection in Section \ref{sec:envelope} and define, as precisely as possible, what constitutes a temporal envelope.

@@

We next explore, in Section \ref{sec:curvature}, the concept of discrete curvature, that will be used in Section \ref{sec:envelopedetection}, where the envelope detection algorithm is presented.

Section \ref{sec:quality} discusses the difficulties in assessing the quality of an envelope, proposing qualitative and quantitative approaches, and demonstrates that the method presented conforms to four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.

In Section \ref{sec:comparison} we compare the presented algorithm with traditional envelope detection methods in terms of accuracy and time and present, in Section \ref{sec:extensions}, potential applications of the algorithm, some of which might not be immediately obvious.

Links to interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, which also contains an implementation of the method in pure Python.
A Python module can be installed from the Python Package Index (PyPI) with the command “pip install signal-envelope”. Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.

@@

The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to artificial waves, often with well-defined frequencies \cite{2011TurnerDemodulation}.

While there is no general formal definition for the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the consensus is that an envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}.

In this context, the most mathematically sound definition of an envelope, as introduced by Gabor in 1946 \cite{2007HahnHistory}, involves representing the underlying wave as an analytic signal. In his work, \cite{1946GaborTheory} applies the then relatively new mathematical machinery of quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform could be applied to a real signal in order to obtain an equivalent complex signal, later known as the analytic signal.

This analytic signal has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \cite{2016HePraat} where $ S(t) $ is the original real signal. $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $. This representation is common in classical AM-FM literature.

@@

Fig. \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise. It is readily noticeable from Fig. \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected in the envelope.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of degree 3, as obtained by the Hilbert Transform approach. In the first half of the samples, the sinusoid is free of noise, while in the second half Gaussian noise, with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave, was added to the base signal.}
  \label{fig:AnalyticSignal}
\end{figure}

For real-world signals, such as an alto singer sustaining a steady note as shown in Fig. \ref{fig:AnalyticSignalAlto}, the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave. With further processing, however, such as filtering, a smoother output can be obtained.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignalAlto.pdf}
  \caption{A section of the representation of an alto singer sustaining a steady note comparing the envelopes obtained by the Hilbert transform with and without further filtering.}
  \label{fig:AnalyticSignalAlto}
\end{figure}

In general, the problem of envelope detection, or demodulation, can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two components $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \cite{2011TurnerDemodulation}.

@@

We address this problem by assuming $ C(t) $ to be normalized between $ \{-1, 1\}$ or, in other words $ |\max(C(t))| = 1 $. This assumption does not cause any loss of generality since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.

In this work we are concerned with the discrete version of the envelope detection problem: given a finite digital wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, obtaining the temporal envelope $ \mathbf{e} $ of $ \mathbf{w} $.

The preceding definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals. In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each sample of the discrete signal, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second. We can thus define:

@@

The $ \odot $ operator in Equation \ref{eq:Envelope} stands for the Hadamard product, denoting element-wise multiplication of two vectors. Fig. \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by multiplying the normalized carrier wave \textbf{c} using the envelope \textbf{e}.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} composed by the element-wise multiplication of a previously known envelope \textbf{e} and carrier \textbf{c}. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

In Fig. \ref{fig:Envelope} a signal arising from the multiplication of a known carrier wave \textbf{c} and a known envelope \textbf{e} is shown; this helps to illustrate a key concept: note from Fig. \ref{fig:Envelope} that, only at some local extrema, the envelope can be inferred without prior knowledge of the underlying carrier, being, in fact, equal to the value of the instantaneous amplitude of the wave $ w_i $. That is, $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.

@@

Note that, in Fig. \ref{fig:Envelope}, the carrier, the original signal and even the envelope are representations of similar entities: signals in the same coordinate system. This is in line with \cite{2011CaetanoImproved,2011TurnerDemodulation,2017LyonsDigital} but in contrast with some representations found in the literature of amplitude and frequency modulation of narrowband signals \cite[e.g.,][]{1996Loughlinamplitude} where the FM and AM of a signal are represented in different coordinate systems, commonly frequency versus time and amplitude versus time, respectively.

@@

For this reason, we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_{i_j} \lvert) $, that is, $ \lvert w_{i_j} \lvert $, the absolute value of the extrema of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa.

This relation is illustrated in Fig. \ref{fig:Pulses}. We call it P to emphasize the connection with the pulses of \textbf{w}.
While the exact relation between $ n $, the number of samples of the original signal, and $ m $, the cardinality of the set of points, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals. In practice, $ m \ll n $ for most real-world signals.

@@

Geometrically, we can think of these points as points in a vector space on a non-standard, orthogonal basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (a, 0), \mathbf{b}_y = (0, i) $ where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value. $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.

We are interested in representing the point in the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space, making the geometric representation consistent.

To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively canceling the unit of the vertical components of the points. If we proceeded similarly concerning the horizontal components,
the direct link between P and \textbf{w}, that arises from the fact that the abscissas of the points in P are numerically equivalent to the indices $ i $ of \textbf{w}, would be lost.

It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior. In this way, we can consider both axes as having unit $ i $, and the relationship is preserved, making it easier to recover the envelope points in the original coordinates later.

Considering $P_{\mathbf{b}}$ as a point of the set P represented on the original basis and $P_{\mathbf{s}}$ the same point as represented on the standard basis, we can formalize this transformation as in Equation \ref{eq:Basis}:

\begin{align}\label{eq:Basis}
\begin{split}
P_{\mathbf{b}} &= (x_1, x_2) \\
\beta &= \frac{\left( \frac{i_{j-1} - i_j}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
P_{\mathbf{s}} &= (x_1, \beta x_2) \quad \forall \  P \in \text{P} \\ 
\end{split}
\end{align}

@@

Algorithms such as the convex hull and the alpha shapes solve that problem by discriminating which points of a set are part of its frontier.

@@

Alternatively, one can picture a circle of fixed radius being rolled above the set P, and marking the points it touches as part of the envelope.

@@

The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \cite{2001CoeurjollyDiscrete}. Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity, as is generally not the case for approaches that involve curve fitting.

Concerning direct methods \cite{2014CarrollSurvey}, derive three such definitions based on the approximation of a circle by an inscribed, centered, and circumscribed polygon. In the context of three-dimensional meshes, \cite{2016VasaMultivariate} evaluate a range of existing estimators from a multivariate point of view.
Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case). Since we are interested in the change of direction, what the term curvature ultimately means, between two adjacent points, a definition of discrete curvature over the edges of a discrete curve would be preferable.

Thus, the set V of the vectors from each point of P to the next, shown in the Fig. \ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values thus obtained to infer the average curvature of the whole discrete curve.

@@

We formalize the set V of vectors in Equation \ref{eq:Vectors} below.

@@

One of those edges, in the form of the vector $ v_k \in \text{V} $, is singled out in Fig. \ref{fig:DiscreteCurvature}. $ v_k $ represents the net displacement of the envelope from the extrema $ P_{j} = (x_j, y_j) $ to $ P_{j+1} = (x_{j+1}, y_{j+1}) $.

We assume from $ v_k $ that, near the vector, the envelope defines a vertical displacement of $ v_{k,y} $ in the horizontal distance of $ v_{k,x} $; since we don't yet know which points of P are part of the envelope, we can't be certain that the envelope passes through $ P_{j-1} $ before reaching $ P_{j} $, or even if it reaches $ P_{j} $.

Nevertheless, assuming it passes through $ P_{j} $, and is limited to the diplacement observed, the envelope might have passed anywhere in the line segment from points $ (x_j - (x_{j+1} - x_j), y_j - (y_{j+1} - y_j)) $ to $ (x_j - (x_{j+1} - x_j), y_j + (y_{j+1} - y_j)) $, the bold line segment before $ v_k $ in Fig. \ref{fig:DiscreteCurvature}.

@@

The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose unit tangent performs the same change in direction, in the same horizontal distance, as illustrated in Fig. \ref{fig:DiscreteCurvature}. The average curvature of P will be obtained by the average of all radii $ r_k $.

From Fig. \ref{fig:DiscreteCurvature} one can see that $ r_k = v_{k,x} / \sin(\theta_k) $ where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of the horizontal line and the slope $ m_k = v_{k,y} / v_{k,x} $ defined by $ v_k $ using the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.

@@

The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Equation \ref{eq:Radius}. This equation has the desirable property that $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.

The radius is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally far from each other.

>>>>

\begin{abstract} 
Despite being an elusive concept, the temporal amplitude envelope of a signal is essential for its complete characterization, being the primary information-carrying medium in spoken voice and telecommunications, for example. Envelope detection techniques have applications in medicine, sound classification, sound synthesis, seismology, and speech recognition. Nevertheless, a general approach to envelope detection doesn't exist. Most methods involve manual intervention, in the form of filter design, smoothing, and other specific design choices based on prior knowledge of the signals under investigation. To address this problem, we propose an algorithm that uses characteristics of a signal to estimate its envelope, eliminating the necessity of parameter tuning. The method draws inspiration from geometric concepts: the discrete signal is mapped to a set of points in the Cartesian plane, where a new measure of discrete curvature is used to identify the samples that are part of the envelope. Due to its geometric nature, the method can optionally generate the superior and inferior envelopes of a given wave. The algorithm compares favorably with classic envelope detection techniques based on smoothing, filtering, and the Hilbert Transform, besides being physically plausible. We provide visualizations of the envelope extracted for various real-world signals with diverse characteristics, such as voice, spoken and sang, and pitched and unpitched musical instruments, and discuss some approaches to objectively assess the quality of the obtained envelopes. An optimized implementation is available via the Python Package Index (PyPI), while interactive visualizations and source code are available from the accompanying GitHub repository.
\end{abstract}

@@

\section{Introduction}
Envelope detection, also known as demodulation, is ubiquitous in analog and digital signal processing (DSP) \cite{2011CaetanoImproved}. Nevertheless, the literature in this area is fragmented \cite{2017LyonsDigital}, with most envelope detection techniques being designed to account for particular settings involving narrowband signals with moderate noise content.

Those constraints exclude most physical signals, such recorded sounds, for example; this problem arises in part due to the lack of a strict mathematical definition of a temporal envelope \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}.

@@

The temporal envelope also helps to impart emotion and identity to the human voice \cite{2018ZhuContributions} and music in general: Concert halls with envelope preserving characteristics are deemed more pleasant \cite{2011LokkiEngaging}.

Specially when dealing with broadband signals, approaches tailored to specific applications, exploiting prior knowledge of the waves, are prevalent, such as the one presented by \cite{2014YangFast} for the distributed monitoring of fiber optic, or the one formulated by \cite{2018AssefModeling} in the context of medical ultrasound imaging.

The problem of envelope detection of a discrete signal, traditionally associated with the field of digital signal processing, can be made equivalent to the geometric problem of defining the shape of a set of points in $ \mathbb{R}^2 $.

Considering the standard graphic representation of a discrete wave, where the horizontal axis represents time and the vertical axis represents amplitude, one can think of each sample as a point in a plane. Provided a coherent mapping from this space to the Cartesian coordinate system is available, the original DSP problem can be transformed into the geometric problem of finding the shape of this set of points.

One direct way of defining the shape of a set of points in $ \mathbb{R}^2 $ is by using the concept of a convex hull, as defined in \cite{2008BergComputational}. The convex hull of a set of points $ S $ in two-dimensional space is the smallest convex subset of the plane that contains all the points in $ S $.

To get an intuition, one can think of the points in a set as nails in a wooden board, around which a rubber band is stretched. The polygon defined by the rubber band is the convex hull of this set.
Convex hulls are unique, in the sense that a given set of points define one and only one convex hull.

@@

Fig. \ref{fig:AlphaShape} illustrates an alpha shape, for $ \alpha=10 $, of a set of points. Note that two points of the set lie on the circle illustrated, but the open disc it defines contains no points of the set. An analogous construction is possible for every point at the frontier.

Conversely, any circle of radius $ \alpha=10 $ containing at least one of the interior points defines an open disc that also contains at least another point of the set P.

@@

Once a radius $ \alpha $ is defined, alpha shapes exhibit the uniqueness property: for a given set of points and a given radius, only one alpha shape exists. When the radius $ \alpha $ of the circle tends to infinity, the alpha shape of the set tends to its convex hull.

Consider, for example, the artificially generated discrete wave in Fig. \ref{fig:Hulls}, where the links between consecutive samples were de-emphasized to highlight the “set of points” nature of a signal: two alpha shapes, with different values of $ \alpha $, and the convex hull are illustrated. As $ \alpha $ increases, the contour of the set of points becomes coarser, and closer to the convex hull.

@@

Alpha shapes are used in areas such as detection of features in images \cite{2016VarytimidisAlpha}, reconstruction of surfaces from clouds of points \cite{2015WuAutomated} and spectroscopy \cite{2019XuModeling}.

@@

Other steps in the direction of applying geometric algorithms to envelope detection were made by \cite{2015YangSkeleton}, with an algorithm based on the construction of a skeleton underlying the digital wave of interest, and by \cite{2015YangRepresenting}, via the direct translation of computer vision methods to the task of envelope detection.

@@

We start with a brief background of envelope detection in Section \ref{sec:envelope} where we define, as precisely as possible, what constitutes a temporal envelope.

@@

We next explore, in Section \ref{sec:curvature}, the concept of discrete curvature, used in Section \ref{sec:envelopedetection} to discriminate which samples of a discrete signal belong to the envelope.

Section \ref{sec:quality} discusses the difficulties in assessing the quality of an envelope, suggesting qualitative and quantitative approaches. It is also demonstrated that the method presented conforms to the four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}. We also compare the presented algorithm with traditional envelope detection methods in terms of accuracy and processing time.

Section \ref{sec:extensions} discusses potential applications of the algorithm, some of which might not be immediately obvious.

Links to interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, which also contains an implementation of the method in pure Python.

A Python module can be installed from the Python Package Index (PyPI) with the command “pip install signal-envelope”. Those working under Windows 64-bit can benefit from a specialized dynamic library implemented in C++, automatically used within this module. This library can be obtained separately via the repository's release section.

@@

The interest in the theory of envelope detection arose in the analog domain, with the widespread adoption of radio communications, where the problem is typically restricted to artificial waves with well-defined frequencies \cite{2011TurnerDemodulation}.

While there is no general formal definition for the envelope of broadband signals \cite{2014Yangnovel,2015Yangtheoretical,2019Jiaempirical}, the intuition is that the ideal envelope is a smooth curve situated as close as possible to the signal \cite{2019Jiaempirical}, without intersecting with it.

The most mathematically sound definition of an envelope, introduced by Gabor in 1946 \cite{2007HahnHistory}, involves representing the underlying wave as an analytic signal. In his work \cite{1946GaborTheory}, Gabor employs the then relatively new mathematical machinery of quantum mechanics to unify the time and frequency-domain representations of a wave, showing how the Hilbert transform can be applied to a real signal in order to create an equivalent complex signal.

This complex signal became known as the analytic signal, and has the form $ A(t) = S(t) + \mathbb{H}(S(t)) \ \text{i} $ \cite{2016HePraat}, where $ S(t) $ is the original real signal. $ \mathbb{H}(S(t)) $, the Hilbert transform of the original signal, becomes the imaginary part of the analytic signal; the envelope of a signal thus represented can be straightforwardly obtained by the computation of the complex modulus of $ \mathbb{H}(S(t)) $. The analytic signal representation is common in traditional AM-FM literature.

@@

Fig. \ref{fig:AnalyticSignal} exemplifies the Hilbert envelope, for a pure sinusoid modulated by a polynomial, illustrating the difference in the smoothness of the envelope in the presence and absence of Gaussian white noise. It is readily noticeable from Fig. \ref{fig:AnalyticSignal} that, as soon as noise is introduced in the original signal, it is directly reflected on the envelope.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignal.pdf}
  \caption{Envelope of a pure sinusoid with a local frequency of 20 cycles, modulated by a polynomial of degree 3, obtained by the Hilbert Transform approach. In the first half of the samples, the sinusoid is free of noise. In the second half Gaussian noise with a standard deviation of $ 1/10 $ of the maximum amplitude of the wave was added to the signal.}
  \label{fig:AnalyticSignal}
\end{figure}

For real-world signals, such the one of an alto singer sustaining a steady note partially shown in Fig. \ref{fig:AnalyticSignalAlto}, the Hilbert transform alone provides a poor envelope, since it retains most of the frequencies of the underlying wave. With further processing, such as filtering, a smoother output can be obtained.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/01AnalyticSignalAlto.pdf}
  \caption{Comparison of the envelopes obtained by the Hilbert transform with and without further filtering, for a section of the representation of an alto singer sustaining a steady note.}
  \label{fig:AnalyticSignalAlto}
\end{figure}

In general, the problem of envelope detection can be interpreted as the task of, given a continuous wave $ W(t) $, decomposing it in two waves $ E(t) $ and $ C(t) $, such that $ W(t) = E(t) C(t) $ \cite{2011TurnerDemodulation}.

@@

We address this problem by assuming $ C(t) $ to be normalized between $ \{-1, 1\}$ such that $ |\max(C(t))| = 1 $. This assumption does not cause any loss of generality, since, given an arbitrary $ C(t) $, we can obtain its normalization by dividing the function by its absolute global maximum, that is, $ \hat{C}(t) = C(t) / \max(|C(t)|) $, provided that $ \max(|C(t)|) \ne 0 $.

In this work we are concerned with the discrete version of the envelope detection problem: given a finite discrete wave represented by the vector $ \mathbf{w} \in \mathbb{R}^n $, obtaining the temporal envelope $ \mathbf{e} \in \mathbb{R}^n $ of $ \mathbf{w} $.

The continuous definitions can be translated to this discrete scenario assuming that the discrete quantities arise from observing the continuous ones at regular time intervals. In that case, the equality $ i = t \ \text{fps} $ can be used to link both settings, where $ i $ is the index of each sample of the discrete signal, $ t $ stands for the time in seconds, and fps is the frame rate, or the number of observations made in one second. We can thus define:

@@

The $ \odot $ operator in Equation \ref{eq:Envelope} stands for the Hadamard product and denotes element-wise multiplication of two vectors. Fig. \ref{fig:Envelope} provides an example of the vectors just defined: \textbf{w} is obtained by element-wise multiplication between the normalized carrier wave \textbf{c} and the envelope \textbf{e}.

\begin{figure}[ht!]
  \centering
    \includegraphics{images/02Envelope.pdf}
  \caption{Example of a discrete wave \textbf{w} arising from the element-wise multiplication of an envelope \textbf{e} and an carrier \textbf{c}, both previously known. The local extrema of \textbf{w} are highlighted with a circle.}
  \label{fig:Envelope}
\end{figure}

Fig. \ref{fig:Envelope} helps to illustrate a key concept: only at some local extrema the envelope can be inferred without prior knowledge of the underlying carrier.

TODO
% being, in fact, equal to the value of the instantaneous amplitude of the wave $ w_i $. That is, $ c_i = 1 $ at those points, rendering $ e_i = 1 w_i $.

@@

Note that, in Fig. \ref{fig:Envelope}, the carrier, the original signal and even the envelope are representations of similar entities: signals in the same coordinate system. This is in line with \cite{2011CaetanoImproved,2011TurnerDemodulation,2017LyonsDigital}, but in contrast with some representations found in the literature of amplitude modulation of narrowband signals \cite[e.g.][]{1996Loughlinamplitude}, where the FM and AM of a signal are represented in different coordinate systems, commonly frequency versus time and amplitude versus time, respectively.

@@

For this reason, we define P as the set of the points $ \{P_0, P_1, \cdots, P_{m-1}\} $ where $ P_j = (i_j, \lvert w_{i_j} \lvert) $, that is, $ \lvert w_{i_j} \lvert $, the absolute value of the extremum of each pulse, becomes the ordinate of each point and $ i_j $, its original index, becomes the abscissa. This is illustrated in Fig. \ref{fig:Pulses}.

While the exact relationship between $ n $, the number of samples of the original signal, and $ m $, the cardinality of the set P, is dependent on factors such as the frequency and sampling rate of the original wave, the inequality $ m \le n $ holds for all discrete signals. In practice, $ m \ll n $ is true for most real-world signals.

@@

Geometrically, we can think of these points as points in a vector space on a non-standard, orthogonal basis vector $ \mathbf{b} = \{ \mathbf{b}_x, \mathbf{b}_y \}$ with $ \mathbf{b}_x = (i, 0), \mathbf{b}_y = (0, a) $, where $ a $ is an amplitude unit such as decibel, a measure of the instantaneous sound pressure, volt or even a fraction of a fixed maximum value. $ i $ represents the index of the sample, and is linked to time by the aforementioned relation $ i = t \ \text{fps} $.

We are interested in representing those points on the Cartesian coordinate system, and for that we need a transformation that places those points on the standard basis $ \mathbf{s} = \{ \mathbf{s}_x, \mathbf{s}_y \} = \{ (1,0),(0,1) \}$ of this vector space.

To achieve that, we divide the ordinate of each point by the average of all ordinates, effectively canceling the unit of the vertical components of the points. If we were to proceed similarly concerning the horizontal components,
the direct link between P and \textbf{w} would be lost.

It is best to leave the abscissas intact and multiply the now unitless vertical coordinates of each point by the average of the difference between the abscissa of a point and the abscissa of the immediately posterior point.

In this way, we can consider both axes as having unit $ i $. Dividing them by $ i $, we obtain the effect of representing the points on the standard basis. Numerically, the relationship is preserved, since the abscissas of the points $ P_{\mathbf{s}} $ are numerically equivalent to the indices $ i $ of \textbf{w}, making it straightforward to recover the envelope points on the original coordinates later.

Considering $ P_{\mathbf{b}} $, a point of the set P represented on the original coordinate system and $ P_{\mathbf{s}} $ the same point represented on the Cartesian coordinate system, we formalize this transformation in Equation \ref{eq:Basis}:

\begin{align}\label{eq:Basis}
\begin{split}
P & = (x, y) \\
\beta &= \frac{\left( \frac{i_{j+1} - i_j}{m-1} \right)}{\left( \frac{\sum_{j=0}^{m-1} \lvert w_j \lvert}{m} \right)}  \\
P_{\mathbf{b}} &= x \mathbf{b}_x + y \mathbf{b}_y = x (i, 0) + y (0, a) = (x i, y a) \\
P_{\mathbf{s}} &= x \mathbf{s}_x + \beta y \mathbf{s}_y = x (1, 0) + \beta y (0, 1) = (x, \beta y) \\
&\forall \  P \in \text{P} \\ 
\end{split}
\end{align}

@@

Algorithms, such as the convex hull and the alpha shapes, solve this problem by discriminating which points of a set are part of its frontier.

@@

Alternatively, one can picture a circle of fixed radius being rolled above the points in the set P, marking the points it touches as part of the envelope.

@@

The two general approaches are the derivation of direct methods that use characteristics of the discrete curve to calculate the curvature, or the use of the curvature of a smooth, continuous curve fitted to the original discrete curve \cite{2001CoeurjollyDiscrete}. Since we are dealing with a potentially large set of points, it is important for this measure to act locally, to assure $ O(n) $ complexity; this is generally not the case for approaches involving curve fitting.

Concerning direct methods, \cite{2014CarrollSurvey} presents three such definitions based on the approximation of a circle by an inscribed, centered, and circumscribed polygon. In the context of three-dimensional meshes, \cite{2016VasaMultivariate} evaluate a range of existing estimators, from a multivariate point of view.

Those approaches define the discrete curvature in relation to the vertices of a discrete curve (or mesh, in the three-dimensional case). Since we are interested in the change of direction, what the term curvature ultimately means, between two adjacent points, a definition of discrete curvature over the edges of a discrete curve would be preferable.

Thus, the set V of the vectors from each point of P to the next, shown in the Fig. \ref{fig:Vectors}, will be used to estimate the average curvature of P: We will calculate the equivalent circle for each vector, and average the values obtained to infer the average curvature of the discrete curve.

@@

We formalize the set V of vectors in Equation \ref{eq:Vectors}.

@@

One of those edges, in the form of the vector $ v_k \in \text{V} $, is singled out in Fig. \ref{fig:DiscreteCurvature}. The vector $ v_k $ represents the envelope's displacement from the extrema $ P_{j} = (x_j, y_j) $ to $ P_{j+1} = (x_{j+1}, y_{j+1}) $.

We assume that, near the vector $ v_k $, the envelope defines a vertical displacement $ v_{k,y} = y_{j+1} - y_j $ and a horizontal displacement $ v_{k,x} = x_{j+1} - x_j $. Since we don't yet know which points of P are part of the envelope, we can't be certain that the envelope passes through $ P_{j-1} $ before reaching $ P_{j} $, or even if it reaches $ P_{j} $.

Nevertheless, assuming it passes through $ P_{j} $, and is limited to the displacements just observed, the envelope might have passed anywhere in the line segment from points $ (x_j - (x_{j+1} - x_j), y_j - (y_{j+1} - y_j)) $ to $ (x_j - (x_{j+1} - x_j), y_j + (y_{j+1} - y_j)) $, the bold line segment before $ v_k $ in Fig. \ref{fig:DiscreteCurvature}, before reaching $ P_{j} $.

@@

The rationale then is to find, for each vector $ v_k $, the radius $ r_k $ of the equivalent circle whose unit tangent performs the same change in direction, in the same horizontal distance. The average curvature of P will be obtained by the average of all radii $ r_k $.

From Fig. \ref{fig:DiscreteCurvature} one can see that $ r_k = v_{k,x} / \sin(\theta_k) $, where $ v_{k,x} $ is the component of $ v_k $ in the horizontal direction; $ \theta_k $ itself can be obtained using the slope $ m_0 = 0 $ of the horizontal line and the slope $ m_k = v_{k,y} / v_{k,x} $, defined by $ v_k $, using the equality $ \theta_k = \arctan \big( (m_k - m_0) / (1 + m_k m_0) \big) $.

@@

The radius $ r_k $ of the circle equivalent to each $ v_k $ is defined in Equation \ref{eq:Radius}. This equation has the desirable property $ v_{k,y} \to 0 \implies v_{k,y} \to \infty $, that is, the radius tends to infinity when no vertical oscillation occurs, effectively turning the circle into a line.

The radius is also directly proportional to $ v_{k,x} $, making the radius larger the more the points are horizontally distant from each other.
